RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-11-12 casestudy.tex
Minor changes, typos, improvements

Recall that \textbf{requirement 1} of our RationalGRL framework is that the argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase. To get a sense of such discussions, we performed a case study to see which types of discourse are used during discussions of system requirements, and how these discourse types can be captured as argument schemes and critical questions. We manually coded transcripts of such discussions using a list of argument schemes and critical questions based on GRL and PRAS. In this section we present our case study. All original transcripts, codings, and models are available in our online repository, which can be found at:

@@

Before we started coding the transcripts, we drew up an initial list of 11 argument schemes (AS0-AS10 in Table~\ref{table:transcripts:results:argumentschemes}), representing \emph{claims} about the goal model containing the requirements of the system. AS0 to AS4 are schemes that concern a single element of the goal model. So, for example, AS0 represents the claim `$a$ is a relevant actor for the system', and AS3 represents the claim `$G$ is a goal for the system'. AS5 to AS10 are claims about the links between GRL elements. Our initial list also contained 18 critical questions, inspired by the questions associated with the original Practical Reasoning Argumentation Scheme \cite{atkinson2007}. CQ2 to CQ12 (Table~\ref{table:transcripts:results:argumentschemes}) are examples of these critical questions, other examples are `Is the softgoal legitimate?' and `Are there alternative ways to contribute to the same softgoal?'.

@@

During the coding, new argument schemes and critical questions were added to the list. For example, we found that the discussants often talk about tasks decomposing into sub-tasks, so we added AS10 and CQ10a. Furthermore, because there were many discussions on the relevance and the clarity of the names of elements, two generic critical questions CQ12 and CQ13 were added. The final results of the coding can be found in Table~\ref{table:transcripts:results:argumentschemes}. We found a total of 159 instantiations of argument schemes AS0-AS10. The most used argument scheme was AS2: ``Actor $A$ has task $T$'', however, each argument scheme is found in transcripts at least twice. A large portion (about 60\%) of argument schemes involved discussions about tasks of the information system (AS2, AS10). We coded 41 applications of critical questions. Many critical questions (about 55\%) involved clarifying the name of an element, or discussing its relevance (CQ12, Gen).

Our coding further led us to identify three different operations, that is, different effects an argument or critical question can have on a goal model: an argument can introduce a new goal model element (\textsf{INTRO}); it can disable (i.e. attack) a goal model element (\textsf{DISABLE}); or it can replace a goal model element (\textsf{REPLACE}). Consider, for example, Table~\ref{table:transcripts:traffic-light} in Appendix~\ref{sect:transcripts:excerpts}. First, an argument is posed that introduces a number of tasks. A counterargument is then given against one of these tasks (\emph{add traffic light}), which is subsequently disabled. An example of replacement is given in Table~\ref{table:transcript:decomposition} (Appendix~\ref{sect:transcripts:excerpts}): what used to be an AND-decomposition is changed into an OR-decomposition. We will further discuss these operations in Section~\ref{sect:overview} with various examples, and w formalize the operations in Section~\ref{sect:formalframework} with algorithms.

>>>>

Recall that \textbf{requirement 1} of our RationalGRL framework is that the argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase. To get a sense of such discussions, we performed a case study to examine which types of discourse are used during discussions of system requirements, and how these discourse types can be captured as argument schemes and critical questions. We manually coded transcripts of such discussions using a list of argument schemes and critical questions based on GRL and PRAS. In this section we present our case study. All original transcripts, codings, and models are available in our online repository, which can be found at:

@@

Before we started coding the transcripts, came up with an initial list of 11 argument schemes (AS0-AS10 in Table~\ref{table:transcripts:results:argumentschemes}), representing \emph{claims} about the goal model containing the requirements of the system. We used no particular method to develop this initial list besides our own intuition about which part of a goal model would be likely to be discussed.

AS0 to AS4 are schemes that concern a single element of the goal model. For example, AS0 represents the claim `$a$ is a relevant actor for the system', and AS3 represents the claim `$G$ is a goal for the system'. AS5 to AS10 are claims about the links between GRL elements. Our initial list also contained 18 critical questions, inspired by the questions associated with the original Practical Reasoning Argumentation Scheme \cite{atkinson2007}. CQ2 to CQ12 (Table~\ref{table:transcripts:results:argumentschemes}) are examples of these critical questions, other examples are `Is the softgoal legitimate?' and `Are there alternative ways to contribute to the same softgoal?'.

@@

During the coding, new argument schemes and critical questions were added to the list. For example, we found that the discussants often talk about tasks decomposing into sub-tasks, so we added AS10 and CQ10a. Furthermore, since there were many discussions on the relevance and the clarity of the names of elements, two generic critical questions CQ12 and CQ13 were added. The final results of the coding can be found in Table~\ref{table:transcripts:results:argumentschemes}. We found a total of 159 instantiations of argument schemes AS0-AS10. The most used argument scheme was AS2: ``Actor $A$ has task $T$'', however, each argument scheme is found in transcripts at least twice. A large portion (about 60\%) of argument schemes involved discussions about tasks of the information system (AS2, AS10). We coded 41 applications of critical questions. Many critical questions (about 55\%) involved clarifying the name of an element, or discussing its relevance (CQ12, Gen).

Our coding further led us to identify three different operations, that represent different effects an argument or critical question can have on a goal model: an argument can introduce a new element in the goal model (\textsf{INTRO}); it can disable (i.e. attack) a goal model element (\textsf{DISABLE}); or it can replace an element in the goal model with another one (\textsf{REPLACE}). Consider, for example, Table~\ref{table:transcripts:traffic-light} in Appendix~\ref{sect:transcripts:excerpts}. First, an argument is posed that introduces a number of tasks. A counterargument is then given against one of these tasks (\emph{add traffic light}), which is subsequently disabled. An example of replacement is given in Table~\ref{table:transcript:decomposition} (Appendix~\ref{sect:transcripts:excerpts}): what used to be an AND-decomposition is changed into an OR-decomposition. We will further discuss these operations in Section~\ref{sect:overview} with various examples, and w formalize the operations in Section~\ref{sect:formalframework} with algorithms.
